---
title: "A quick tour of ppgmmga"
author: 
- Alessio Serafini, Luca Scrucca
date: "`r format(Sys.time(), '%d %b %Y')`"
output: 
  rmarkdown::html_vignette:
    toc: true
    number_sections: false
    # highlight: "monochrome"
    css: "vignette.css"
vignette: >
  %\VignetteIndexEntry{A quick tour of ppgmmga}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
library(knitr)
opts_chunk$set(fig.align = "center", 
               out.width = "90%",
               fig.width = 6, fig.height = 5,
               dev.args=list(pointsize=10),
               par = TRUE, # needed for setting hook 
               collapse = TRUE, # collapse input & ouput code in chunks
               warning = FALSE)

knit_hooks$set(par = function(before, options, envir)
  { if(before && options$fig.show != "none") 
       par(family = "sans", mar=c(4.1,4.1,1.1,1.1), mgp=c(3,1,0), tcl=-0.5)
})
```

# Introduction 

An R package implementing a Projection Pursuit algorithm based on finite Gaussian Mixtures Models for density estimation using Genetic Algorithms (PPGMMGA) to maximise an approximated Negentropy index. The **ppgmmga** algorithm provides a method to visualise high-dimensional data in a lower-dimensional space, with special reference to reveal clustering structures.

```{r, message = FALSE, echo=1}
library(ppgmmga)
cat(ppgmmga:::ppgmmgaStartupMessage(), sep="")
```

# Banknote data

```{r, fig.width = 6, fig.height = 6}
library(mclust)
data("banknote")
X <- banknote[,-1]
Class <- banknote$Status
table(Class)
clPairs(X, classification = Class)
```

# 1-dimensional ppgmmga

```{r}
pp1D <- ppgmmga(data = X, d = 1, approx = "UT", seed = 1)
pp1D
summary(pp1D)
```

```{r, out.width="60%", fig.width=6, fig.height=4}
plot(pp1D)
```

```{r, out.width="70%", fig.width=7, fig.height=4}
plot(pp1D, class = Class)
```


# 2-dimensional ppgmmga

```{r}
pp2D <- ppgmmga(data = X, d = 2, approx = "UT", seed = 1)
summary(pp2D, check = TRUE)
summary(pp2D$GMM)
```

```{r}
plot(pp2D$GA)
```

```{r}
plot(pp2D)
```

```{r, fig.width = 7, fig.height = 5}
plot(pp2D, class = Class, drawAxis = FALSE)
```

# 3-dimensional ppgmmga

```{r}
gmm <- densityMclust(data = scale(X, center = TRUE, scale = FALSE), G = 2)
pp3D <- ppgmmga(data = X, d = 3, 
                center = TRUE, scale = FALSE, gmm = gmm, 
                gatype = "gaisl", 
                options = ppgmmga.options(numIslands = 2),
                seed = 1)
summary(pp3D$GA)
```

```{r}
plot(pp3D$GA)
```

```{r, fig.width = 6, fig.height = 6}
plot(pp3D)
```

```{r, fig.width = 6, fig.height = 6}
plot(pp3D, class = Class)
```

```{r}
plot(pp3D, dim = c(1,2))
```

```{r, fig.width = 7, fig.height = 5}
plot(pp3D, dim = c(1,3), class = Class)
```
<br>

# References

Scrucca, L. and Serafini, A. (2018) Projection pursuit based on Gaussian mixtures and evolutionary algorithms. Under review.

----

```{r}
devtools::session_info()
```
